% bFaaaP arXiv preprint --- full revision (2026-06-14)
% Compiles with: pdflatex -> bibtex -> pdflatex -> pdflatex  (uses refs.bib).
% Suggested arXiv categories: primary cs.HC, cross-list cs.RO (and eess.SY).
% The paper body (Introduction -> Conclusion) is shared with the ACM TACCESS
% version via draft_sections/body.tex, so both stay in sync.
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage[hidelinks]{hyperref}
\usepackage{authblk}
\usepackage{caption}

% The shared body uses \Description{} (figure alt-text) for the ACM/acmart
% version; in article it is a harmless no-op.
\providecommand{\Description}[1]{}

\title{\textbf{bFaaaP: An Inclusive, Head-Angle Piano-Pedal Interaction that
Quantitatively Reproduces a Pianist's Intended Pedalling---Foot-Free, for
Acoustic and Electronic Pianos}}

\author[1,2]{Tomoyuki Shishido\thanks{Corresponding author and first author.
ORCID: \href{https://orcid.org/0000-0002-8944-2088}{0000-0002-8944-2088}
(\url{https://orcid.org/0000-0002-8944-2088}). Email:
\texttt{shishid@saaipf.com}.}}
\author[3]{Masahiro Ootaki}
\author[4]{Hiroyuki Narusawa}
\affil[1]{Shishido \& Associates, Horinouchi, Hachioji, 192-0355, Tokyo, Japan}
\affil[2]{SeemeData Labs, Inc., 2-19-2 3F Nakano, Nakano-ku, 164-0001, Tokyo, Japan}
\affil[3]{Ootaki \& Architects, Fujimi 1-19-15, Sayama, 350-1306, Saitama, Japan}
\affil[4]{Naru Science Software, Fujimi-shi, 354-0024, Saitama, Japan}

\date{Preprint --- July 2026}

\begin{document}
\maketitle

\begin{abstract}
Expressive piano performance depends on the sustain (damper) pedal, operated by
foot---excluding players who cannot readily use their feet: wheelchair users and
others with lower-limb impairments, small children, and some elderly or disabled
players. We present \textbf{bFaaaP} (\emph{barrier-Free assist as a Pedal}), an
inclusive, foot-free interaction that operates the pedal from the angle of the
player's \emph{head}: a smartphone tracks head pose with on-device augmented-reality
(AR) face tracking and streams a compact command over Bluetooth Low Energy (BLE) to
a pedal device. Supported by patent examination, our central claim is \emph{not} the
head$\rightarrow$pedal architecture (anticipated by prior art) but a quantitative,
user-tunable control law---the patentable ``key'' to a natural, expressive result:
the player presets a small angular dead-zone (offset $3$--$10^\circ$) and a
multiplier ($10$--$50$), which together fix a secondary, pre-adjustable response
speed that reproduces the pianist's intended pedalling. An engineering trick
decouples the fast AR sampling from the slower BLE rate. Two co-equal realizations
share one controller: a non-destructive robotic actuator for acoustic pianos
(\emph{Pro}), anchored by a pneumatic ``airback'' (our coined term for an inflatable
air-braced anchor) that absorbs the reaction force without modifying the instrument;
and an electronic sustain switch for digital pianos (\emph{Switch}). In a
human-subject Auxiliary Pedal Effect Evaluation (APEE) with 15 participants, bFaaaP
significantly increased sustained-tone energy ($p<0.01$) and was statistically
indistinguishable from a player's own foot ($p>0.05$), with no significant difference
across classes; one participant with a leg disability and a tracheostomy performed
successfully. With nothing worn on the face and fast setup, bFaaaP has run in formal
public concerts (2018--2025). We release the full hardware and software as open
source.
\end{abstract}

\noindent\textbf{Keywords:} assistive technology; accessibility; inclusive design;
human--machine interaction; foot-free interaction; head-gesture interface;
robotic actuation; accessible musical instruments; clinical evaluation;
open hardware; Bluetooth Low Energy.

% =====================================================================
\input{draft_sections/body}

% =====================================================================
\section*{Acknowledgements}
We thank Yasuhiro Ono for his advice on AI programming, and Taguchi for
contributing the KiCad schematic and electronics of the Pro device. We are grateful to the
Institute of Science Tokyo and to the owner couple of MUSICASA
(\url{https://www.musicasa.co.jp}) for kindly providing venues for our
demonstrations and concerts. We sincerely thank the participants of the Auxiliary
Pedal Effect Evaluation (APEE)---the adults, children, and players with
disabilities, and their families---for generously taking part and sharing their
feedback. We also thank the bFaaaP and Platanus members and
collaborators, and the performers and audiences of the demonstration events.
Caricature illustrations of the team are by Saki Shiokawa.

\section*{Funding}
This bFaaaP project was supported by Shishido \& Associates.

\section*{Ethics}
The Auxiliary Pedal Effect Evaluation (APEE) involved human participants: adults,
children aged 7--11, and people with disabilities. Participation was voluntary, and
informed consent was obtained in writing for every participant. Adult participants gave
their own written informed consent. For the children, a parent or guardian confirmed
written informed consent (arranged through a piano teacher who assisted the study), and
the child signed the consent form themselves. Participants with disabilities took part
with the written consent of a parent or guardian, who also accompanied them and provided
transport to and from the sessions. Any participant could stop or withdraw at any time,
and the tasks (playing short passages on a piano) posed no more than minimal risk. The
authors are independent practitioners and are not affiliated with an institution that
operates an institutional review board (IRB); no formal IRB review or approval number is
therefore available. The study was nonetheless conducted following the ethical principles
of the ACM Publications Policy on Research Involving Human Participants and
Subjects---minimizing potential harm and protecting participants' privacy and right to
self-determination. All reported data are anonymized (Appendix~A), and no personally
identifying information is included.

\section*{Conflict of interest/Competing interests}
T.S.\ is a chief technology officer at SeemeData Labs, Inc.\ and chief Japanese
patent attorney at Shishido \& Associates. T.S., M.O., and H.N.\ are among the
inventors on the patents underlying bFaaaP (JP~6726319 and JP~7004771;
PCT~WO~2019/176164), which are held by T.~Shishido and Ootaki \& Architects, and
are involved in its commercialization (the \emph{bFaaaPSwitch} iOS application and
the build-to-order pedal device); \emph{bFaaaP} is a registered trademark. The
authors otherwise declare no competing interests.

\section*{Patent availability}
The control method is protected by JP~6726319~B2 and JP~7004771~B2 (priority basis:
PCT~WO~2019/176164~\cite{pct2019}). As a matter of policy the authors intend to license the
patents \emph{free of charge} for genuinely public/open uses and, in particular,
for products or services that enable inclusive social participation by people with
disabilities---even commercial ones. Please contact the corresponding author.

\section*{Data and code availability}
The iOS application source, device firmware, mechanical (CAD/3D-print) designs,
documentation, and the patent file-wrapper materials underlying the APEE data are
openly available in the project repository~\cite{repo}. The intended release
licensing (final decision pending) is a three-layer scheme: \emph{software}
(iOS app and firmware) under the \textbf{Apache License 2.0} (which carries an
explicit patent grant), \emph{hardware} designs (CAD/3D-print/schematic) under
\textbf{CERN-OHL-W-2.0}, and \emph{documentation} under \textbf{CC-BY-4.0};
practising the patented invention in a device is governed by the separate patent
policy above.

% =====================================================================
\input{draft_sections/appendix_apee_data}

% =====================================================================
\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}
